---
title: "Iris_Organising_data"
output: html_document
date: "2024-12-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this notebook, I am trying to merge all the datasets into one from the data_cleaning folder made by Graham. From each small dataset, I try to condense the dataset. Here is how I do it.

In trait_data, I removed "Replicate" "Block""Plot""Range" "Pass" "Hybrid_orig_name" "Hybrid_Parent1""Hybrid_Parent2" .
Then I merge selected_trait_data and env_cov data by Env.
For the weather_season data, there are one row for each day for each Env, so I condensed the weather_season data group by Env, and calculate the average for other rows. Then, I merge the weather dataset with selected_trait_data and env_cov data.
For meta_data, I selected Env,Treatment,City,Farm and merge with other df.
I obtained lat, and long from envirotype_input. I group by Env and calculated the average.
For soil_data, I select only Env, `%K Sat`, `%Ca Sat`, `%Mg Sat`, `%Na Sat`, `% Sand`, `% Silt`, `% Clay` and merge them all together by Env

This is how I got to merged_data5 and obtain the csv. As the file size is too large to put in github. I have uploaded it to our google drive. https://drive.google.com/drive/folders/1pWHllpwmzqtKPwm218Wt96MxhtxkEvhT?usp=drive_link


I would need help on start to fit a very simple model like random forest. Please try it and let me know if there's further adjustments needed in the merged_data5

Another task is to utilize the genomic data, if catherine and graham could continue to work on that.

thank you




```{r}
setwd("~/UIUCDigitalHand")
library(readr)
library(tidyverse)
library(dplyr)
library(lme4)
library(mice)
library(caret)
library(randomForest)
library(glmnet)
```
```{r}
env_cov <- read_csv("data_cleaning/training_data/env_cov.csv")
meta_data <- read_csv("data_cleaning/training_data/meta_data.csv")
soil_data <- read_csv("data_cleaning/training_data/soil_data.csv")
trait_data <- read_csv("data_cleaning/training_data/trait_data.csv")
weather_season <- read_csv("data_cleaning/training_data/weather_season.csv")
weather_year <- read_csv("data_cleaning/training_data/weather_year.csv")
envirotype_input <- read_csv("data_cleaning/output_data/envirotype_input.csv")
```


```{r}
all_df <- c(env_cov,meta_data, soil_data,trait_data, weather_season, weather_year)
```

```{r}
# Print structure of each dataframe with a separator
for(name in names(all_df)) {
    cat("\n=====", name, "=====\n")
    str(all_df[[name]])
}
```

```{r}
# select useful variable in trait_data
# Remove "Replicate" "Block""Plot""Range" "Pass" "Hybrid_orig_name" "Hybrid_Parent1""Hybrid_Parent2"  
selected_trait_data <- trait_data %>%
  select(-Replicate, -Block, -Plot, -Range, -Pass, -Hybrid_orig_name, -Hybrid_Parent1, -Hybrid_Parent2)

# Display the first few rows of the modified dataframe
head(selected_trait_data)
```

```{r}
# Merge selected_trait_data and env_cov data by Env
env_cov
```

```{r}
merged_data <- selected_trait_data %>%
  inner_join(env_cov, by = "Env")

# Display the first few rows of the merged dataframe
head(merged_data)
```

```{r}
summary(merged_data)
```

```{r}
head(weather_season)
```

```{r}
condensed_data <- weather_season %>%
  select(-Date) %>%  # Remove the Date column
  group_by(Env) %>%  # Group by Env
  summarise(across(everything(), mean, na.rm = TRUE))  # Calculate the mean for each group

# Display the first few rows of the condensed dataframe
head(condensed_data)
```

```{r}
# merge merged_data and condensed_data by Env
merged_data2 <- merged_data %>%
  inner_join(condensed_data, by = "Env")

# Display the first few rows of the merged dataframe
head(merged_data2)
```

```{r}
# select useful columns from meta_data
selected_meta_data <- meta_data %>%
  select(Env,Treatment,City,Farm)

# Display the first few rows of the modified dataframe
head(selected_meta_data)
```

```{r}
# merge merged_data2 and selected_meta_data by Env
merged_data3 <- merged_data2 %>%
  inner_join(selected_meta_data, by = "Env")

# Display the first few rows of the merged dataframe
head(merged_data3)
```

```{r}
# put lat and long from the envirotype_input data to the merged_data3
selected_envirotype_input <- 
  envirotype_input %>% 
  select(Env,Latitude, Longitude ) %>%
  group_by(Env) %>%  # Group by Env
  summarise(across(everything(), mean, na.rm = TRUE))  # Calculate the mean for each group

head(selected_envirotype_input)
```

```{r}
# merge merged_data3 and selected_envirotype_input by Env
merged_data4 <- merged_data3 %>%
  inner_join(selected_envirotype_input, by = "Env")

# Display the first few rows of the merged dataframe
head(merged_data4)
```

```{r}
na_counts <- sapply(soil_data, function(x) sum(is.na(x)))

# Display the number of NA values for each column
print(na_counts)
```

```{r}
selected_soil_data <- soil_data %>%
  select(Env, `%K Sat`, `%Ca Sat`, `%Mg Sat`, `%Na Sat`, `% Sand`, `% Silt`, `% Clay`) %>%
  group_by(Env) %>%
  summarise(across(everything(), ~ mean(.x, na.rm = TRUE)))

# Display the first few rows of the resulting dataframe
head(selected_soil_data)
```

```{r}
# merge merged_data4 and selected_soil_data by Env
merged_data5 <- merged_data4 %>%
  inner_join(selected_soil_data, by = "Env")

# Display the first few rows of the merged dataframe
head(merged_data5)
```

```{r}
write_csv(merged_data5, "merged_data5.csv")
```

